---
title: "immobilization calculations"
author: "Dave Costello"
date: "5/24/2021"
output: html_document
---

Function used to calculate phosphorus immobilization rate, time of immobilization, and P factor. See Costello et al. 2022 (https://doi.org/10.1029/2021GB007163) and Supplemental Information for details of calculations. Default values are provided for initial P concentration, initial punch mass, and initial punch C mass, but they likely only apply to the CELLDEX cotton strips deployed in 2015 and the sampling methods described in the manuscript.

```{r function to calculate immobilization}
Pimmob <- function(P.conc.ugg,C.mass,time=21,dd=NA,
                        P.init.ugg = 64,C.mass.init=3.388,mass.init=7.24){
  
  #Set up results
  Pfact = vector('double')
  Pim =vector('double')
  massmax = vector('double')
  maxP = vector('double')
  tmax = vector('integer')
  Pim.dd = vector('double')
  
  for(i in 1:length(P.conc.ugg)){
    
    #First check if there is P data & mass data
    if(is.na(P.conc.ugg[i])|is.na(C.mass[i]))
      {Pfact[i] = NA
      Pim[i] = NA
      Pim.dd[i] = NA
      maxP[i] = NA
      tmax[i] = NA
      massmax[i] = NA}
    
    #Check to confirm there was mass loss
    else if(C.mass[i]>=C.mass.init)
      {Pfact[i] = NA
      Pim[i] = NA
      Pim.dd[i] = NA
      maxP[i] = NA
      tmax[i] = NA
      massmax[i] = NA}
    
      else{
        #Calculate regression parameters of the mass loss vs. P concentration
        #Calculations in Aber and Melillo (using C mass rather than mass loss %)
        slope1 <- (C.mass[i]/C.mass.init-1)/(P.conc.ugg[i]-P.init.ugg)
        inter1 <- 1-slope1*P.init.ugg
        kC = log(C.mass[i]/C.mass.init)/time[i]
          
    #This tests whether intercept > 2 and slope is positive, which indicates that P is being immobilized
        if(inter1<2&slope1<0){
          massmax[i] <- inter1/2 #C mass remaining (%) at max P mass
          maxP[i] <- inter1/(2*-slope1) #P concentration (ug/g) at max P mass
          tmax[i] <- log(massmax[i])/kC #Time to reach max P
          
          #Pfact is the total mass of P (ng) immobilized per punch
          Pfact[i] <- maxP[i]*(massmax[i]*mass.init)-(P.init.ugg*mass.init)
    
  #Pim_rate is the linear rate of P immobilization (ug P/g C/d) until mineralization happens
  Pim[i] <- (maxP[i]-P.init.ugg)*(1/0.42)/tmax[i]
        
  #P immobilization rate can also be calculated per degree day
  dd.d <- dd[i]/time[i]
  Pim.dd[i] = (maxP[i]-P.init.ugg)*(1/0.42)/(tmax[i]*dd.d)
        }
        
  #If mass loss occurred but P mass did not increase then immobilization = 0
    else{Pfact[i] = 0
    Pim[i] = 0
    Pim.dd[i] = 0
    maxP[i] = NA
    tmax[i] = tmax[i]
    massmax[i] = NA}
}
  }

  return(list(
    P.factor=Pfact,
    maxP = maxP,
    massatmax = massmax,
    tmax=tmax,
    Pim.rate = Pim,
    Pim.rate.dd = Pim.dd
    #slope1=slope1,
    #inter1=inter1,
    #decomp = kC
    ))
  
}
```

```{r Examples}
#EXAMPLE AAKO1 strip 4885 (P immobilized) 
Pimmob(P.conc.ugg=273,C.mass=3.32,time=23)

#EXAMPLE 2 AB3 strip 4503 (P concentration decline, no immobilization)
Pimmob(P.conc.ugg=54,C.mass=3.29,time=35)
```

